\begin{thebibliography}{00}

\bibitem{han2016deep}
S. Han, H. Mao, and W. J. Dally, ``Deep compression: Compressing deep neural
networks with pruning, trained quantization and Huffman coding,'' in
\textit{Proc. Int. Conf. Learn. Represent. (ICLR)}, 2016.

\bibitem{jacob2018quant}
B. Jacob \textit{et al.}, ``Quantization and training of neural networks for
efficient integer-arithmetic-only inference,'' in \textit{Proc. IEEE/CVF Conf.
Comput. Vis. Pattern Recognit. (CVPR)}, 2018, pp.\ 2704--2713.

\bibitem{gholami2022survey}
A. Gholami, S. Kim, Z. Dong, Z. Yao, M. W. Mahoney, and K. Keutzer, ``A survey
of quantization methods for efficient neural network inference,'' in
\textit{Low-Power Computer Vision}. Boca Raton, FL, USA: Chapman \& Hall/CRC,
2022, pp.\ 291--326.

\bibitem{schwartz2020green}
R. Schwartz, J. Dodge, N. A. Smith, and O. Etzioni, ``Green AI,''
\textit{Commun. ACM}, vol.\ 63, no.\ 12, pp.\ 54--63, Dec. 2020.

\bibitem{simonyan2014saliency}
K. Simonyan, A. Vedaldi, and A. Zisserman, ``Deep inside convolutional networks:
Visualising image classification models and saliency maps,'' in \textit{Proc.
Int. Conf. Learn. Represent. (ICLR) Workshop}, 2014.

\bibitem{sundararajan2017ig}
M. Sundararajan, A. Taly, and Q. Yan, ``Axiomatic attribution for deep
networks,'' in \textit{Proc. Int. Conf. Mach. Learn. (ICML)}, 2017, pp.\
3319--3328.

\bibitem{selvaraju2017gradcam}
R. R. Selvaraju, M. Cogswell, A. Das, R. Vedantam, D. Parikh, and D. Batra,
``Grad-CAM: Visual explanations from deep networks via gradient-based
localization,'' in \textit{Proc. IEEE Int. Conf. Comput. Vis. (ICCV)}, 2017,
pp.\ 618--626.

\bibitem{bach2015lrp}
S. Bach, A. Binder, G. Montavon, F. Klauschen, K.-R. M\"uller, and W. Samek,
``On pixel-wise explanations for non-linear classifier decisions by layer-wise
relevance propagation,'' \textit{PLoS ONE}, vol.\ 10, no.\ 7, p.\ e0130140,
2015, doi: 10.1371/journal.pone.0130140.

\bibitem{ribeiro2016lime}
M. T. Ribeiro, S. Singh, and C. Guestrin, ```Why should I trust you?':
Explaining the predictions of any classifier,'' in \textit{Proc. ACM SIGKDD Int.
Conf. Knowl. Discovery Data Mining (KDD)}, 2016, pp.\ 1135--1144.

\bibitem{lundberg2017shap}
S. M. Lundberg and S.-I. Lee, ``A unified approach to interpreting model
predictions,'' in \textit{Adv. Neural Inf. Process. Syst. (NeurIPS)}, 2017,
pp.\ 4765--4774.

\bibitem{yeom2021pruning}
S.-K. Yeom \textit{et al.}, ``Pruning by explaining: A novel criterion for deep
neural network pruning,'' \textit{Pattern Recognit.}, vol.\ 115, p.\ 107899,
2021.

\bibitem{becking2022ecq}
D. Becking, M. Dreyer, W. Samek, K. M\"uller, and S. Lapuschkin,
``ECQ$^{\text{x}}$: Explainability-driven quantization for low-bit and sparse
DNNs,'' in \textit{xxAI---Beyond Explainable AI} (Lecture Notes in Computer
Science), vol.\ 13200. Cham, Switzerland: Springer, 2022, pp.\ 271--296, doi:
10.1007/978-3-031-04083-2\_14.

\bibitem{nagel2021whitepaper}
M. Nagel, M. Fournarakis, R. A. Amjad, Y. Bondarenko, M. van Baalen, and T.
Blankevoort, ``A white paper on neural network quantization,'' \textit{arXiv
preprint arXiv:2106.08295}, 2021.

\bibitem{lecun1990obd}
Y. LeCun, J. S. Denker, and S. A. Solla, ``Optimal brain damage,'' in
\textit{Adv. Neural Inf. Process. Syst. (NeurIPS)}, 1990, pp.\ 598--605.

\bibitem{li2017pruningfilters}
H. Li, A. Kadav, I. Durdanovic, H. Samet, and H. P. Graf, ``Pruning filters for
efficient ConvNets,'' in \textit{Proc. Int. Conf. Learn. Represent. (ICLR)},
2017.

\bibitem{frankle2019lottery}
J. Frankle and M. Carbin, ``The lottery ticket hypothesis: Finding sparse,
trainable neural networks,'' in \textit{Proc. Int. Conf. Learn. Represent.
(ICLR)}, 2019.

\bibitem{zeiler2014visualizing}
M. D. Zeiler and R. Fergus, ``Visualizing and understanding convolutional
networks,'' in \textit{Proc. Eur. Conf. Comput. Vis. (ECCV)}, 2014, pp.\
818--833.

\bibitem{adebayo2018sanity}
J. Adebayo, J. Gilmer, M. Muelly, I. Goodfellow, M. Hardt, and B. Kim, ``Sanity
checks for saliency maps,'' in \textit{Adv. Neural Inf. Process. Syst.
(NeurIPS)}, 2018, pp.\ 9505--9515.

\bibitem{ghorbani2019interpretation}
A. Ghorbani, A. Abid, and J. Zou, ``Interpretation of neural networks is
fragile,'' in \textit{Proc. AAAI Conf. Artif. Intell.}, vol.\ 33, 2019, pp.\
3681--3688.

\bibitem{petsiuk2018rise}
V. Petsiuk, A. Das, and K. Saenko, ``RISE: Randomized input sampling for
explanation of black-box models,'' in \textit{Proc. Brit. Mach. Vis. Conf.
(BMVC)}, 2018.

\bibitem{smilkov2017smoothgrad}
D. Smilkov, N. Thorat, B. Kim, F. Vi\'egas, and M. Wattenberg, ``SmoothGrad:
Removing noise by adding noise,'' \textit{arXiv preprint arXiv:1706.03825},
2017.

\bibitem{montavon2018methods}
G. Montavon, W. Samek, and K.-R. M\"uller, ``Methods for interpreting and
understanding deep neural networks,'' \textit{Digit. Signal Process.}, vol.\ 73,
pp.\ 1--15, 2018.

\bibitem{hooker2019benchmark}
S. Hooker, D. Erhan, P.-J. Kindermans, and B. Kim, ``A benchmark for
interpretability methods in deep neural networks,'' in \textit{Adv. Neural Inf.
Process. Syst. (NeurIPS)}, 2019, pp.\ 9737--9748.

\bibitem{kim2018tcav}
B. Kim \textit{et al.}, ``Interpretability beyond feature attribution:
Quantitative testing with concept activation vectors (TCAV),'' in \textit{Proc.
Int. Conf. Mach. Learn. (ICML)}, 2018, pp.\ 2668--2677.

\bibitem{nagel2020adaround}
M. Nagel, R. A. Amjad, M. van Baalen, C. Louizos, and T. Blankevoort, ``Up or
down? Adaptive rounding for post-training quantization,'' in \textit{Proc. Int.
Conf. Mach. Learn. (ICML)}, 2020, pp.\ 7197--7206.

\bibitem{wang2019haq}
K. Wang, Z. Liu, Y. Lin, J. Lin, and S. Han, ``HAQ: Hardware-aware automated
quantization with mixed precision,'' in \textit{Proc. IEEE/CVF Conf. Comput.
Vis. Pattern Recognit. (CVPR)}, 2019, pp.\ 8612--8620.

\bibitem{dong2019hawq}
Z. Dong, Z. Yao, A. Gholami, M. W. Mahoney, and K. Keutzer, ``HAWQ: Hessian
aware quantization of neural networks with mixed-precision,'' in \textit{Proc.
IEEE Int. Conf. Comput. Vis. (ICCV)}, 2019, pp.\ 293--302.

\bibitem{molchanov2017pruning}
P. Molchanov, S. Tyree, T. Karras, T. Aila, and J. Kautz, ``Pruning
convolutional neural networks for resource efficient inference,'' in
\textit{Proc. Int. Conf. Learn. Represent. (ICLR)}, 2017.

\bibitem{blalock2020state}
D. Blalock, J. J. G. Ortiz, J. Frankle, and J. Guttag, ``What is the state of
neural network pruning?'' in \textit{Proc. Mach. Learn. Syst. (MLSys)}, vol.\ 2,
2020, pp.\ 129--146.

\bibitem{samek2021explaining}
W. Samek, G. Montavon, S. Lapuschkin, C. J. Anders, and K.-R. M\"uller,
``Explaining deep neural networks and beyond: A review of methods and
applications,'' \textit{Proc. IEEE}, vol.\ 109, no.\ 3, pp.\ 247--278, Mar.
2021.

\bibitem{he2016resnet}
K. He, X. Zhang, S. Ren, and J. Sun, ``Deep residual learning for image
recognition,'' in \textit{Proc. IEEE Conf. Comput. Vis. Pattern Recognit.
(CVPR)}, 2016, pp.\ 770--778.

\bibitem{stewart1990matrix}
G. W. Stewart and J.-G. Sun, \textit{Matrix Perturbation Theory}. San Diego, CA,
USA: Academic Press, 1990.

\bibitem{samek2017evaluating}
W. Samek, A. Binder, G. Montavon, S. Lapuschkin, and K.-R. M\"uller, ``Evaluating
the visualization of what a deep neural network has learned,'' \textit{IEEE
Trans. Neural Netw. Learn. Syst.}, vol.\ 28, no.\ 11, pp.\ 2660--2673, Nov.\
2017.

\bibitem{yuan2022gnnexplain}
H. Yuan, H. Yu, S. Gui, and S. Ji, ``Explainability in graph neural networks: A
taxonomic survey,'' \textit{IEEE Trans. Pattern Anal. Mach. Intell.}, vol.\ 45,
no.\ 5, pp.\ 5782--5799, May 2023.

\bibitem{dettmers2022llmint8}
T. Dettmers, M. Lewis, Y. Belkada, and L. Zettlemoyer, ``LLM.int8(): 8-bit matrix
multiplication for transformers at scale,'' in \textit{Adv. Neural Inf. Process.
Syst. (NeurIPS)}, 2022, pp.\ 30318--30332.

\bibitem{guo2024lqlora}
H. Guo, P. Greengard, E. P. Xing, and Y. Kim, ``LQ-LoRA: Low-rank plus quantized
matrix decomposition for efficient language model finetuning,'' in \textit{Proc.
Int. Conf. Learn. Represent. (ICLR)}, 2024.

% --- Recently published entries; verified against arXiv (Jul 2026). ---
\bibitem{ranjan2025mixqvit}
N. Ranjan and A. Savakis, ``Mix-QViT: Mixed-precision vision transformer
quantization driven by layer importance and quantization sensitivity,''
\textit{arXiv:2501.06357}, 2025, doi: 10.48550/arXiv.2501.06357.

\bibitem{lin2024quantimit}
S. Park \textit{et al.}, ``Saliency-aware quantized imitation learning for
efficient robotic control,'' \textit{arXiv:2505.15304}, 2025, doi:
10.48550/arXiv.2505.15304.

\bibitem{li2024rdkv}
F. Zuo, Z. Zhou, H. Cong, X. Xi, and H. F. Leung, ``RateQuant: Optimal
mixed-precision KV cache quantization via rate--distortion theory,''
\textit{arXiv:2605.06675}, 2026, doi: 10.48550/arXiv.2605.06675.

\end{thebibliography}
